
```{r, include = FALSE}
leanbuild::set_knitr_image_path()
```

# Using version control with GitHub

## Learning Objectives

```{r, fig.alt="This chapter will demonstrate how to: Understand that git and GitHub are tools that help your analyses be conducted reproducibly and in an open source manner. Create a GitHub account. Set up a GitHub repository for your analyses.", out.width = "100%", echo = FALSE}
leanbuild::include_slide("https://docs.google.com/presentation/d/1IJ_uFxJud7OdIAr6p8ZOzvYs-SGDqa7g4cUHtUld03I/edit#slide=id.g1013f9881e2_0_132")
```
In the introductory part of this course, we discussed [some of the reasons for using GitHub](https://jhudatascience.org/Reproducibility_in_Cancer_Informatics/making-your-project-open-source-with-github.html#github-and-git-allow-you-to) but we didn't get into version control or GitHub's capabilities much beyond its capacity to store code in a place where others can find it.

In this Advanced course, we will dig deeper into Git and GitHub's capabilities so you can really use this to your daily work's advantage. But, to gain the benefit of these deeper GitHub skills it will require some new habits to be formed. Fully embracing the GitHub workflow will make your work more efficient and help you create more transparent and reproducible analyses!




## Get the exercise project files (or continue with the files you used in the previous chapter)

<details> <summary>**Get the Python project example files**</summary>
[Click this link to download](https://raw.githubusercontent.com/jhudsl/Reproducibility_in_Cancer_Informatics/main/chapter-zips/python-heatmap-adv-chapt-4.zip).

```{bash, include = FALSE}
mkdir -p chapter-zips
wget -O chapter-zips/python-heatmap-adv-chapt-4.zip https://raw.githubusercontent.com/jhudsl/Reproducibility_in_Cancer_Informatics/main/chapter-zips/python-heatmap-adv-chapt-4.zip
```

Now double click your chapter zip file to unzip. For Windows you may have to [follow these instructions](https://support.microsoft.com/en-us/windows/zip-and-unzip-files-f6dde0a7-0fec-8294-e1d3-703ed85e7ebc)).

```{bash, include = FALSE}
unzip -o chapter-zips/python-heatmap-adv-chapt-4.zip -d chapter-zips/
```
</details>

<details> <summary>**Get the R project example files**</summary>
[Click this link to download](https://raw.githubusercontent.com/jhudsl/Reproducibility_in_Cancer_Informatics/main/chapter-zips/r-heatmap-adv-chapt-4.zip).

```{bash, include = FALSE}
mkdir -p chapter-zips
wget -O chapter-zips/r-heatmap-adv-chapt-4.zip https://raw.githubusercontent.com/jhudsl/Reproducibility_in_Cancer_Informatics/main/chapter-zips/r-heatmap-adv-chapt-4.zip
```

Now double click your chapter zip file to unzip. For Windows you may have to [follow these instructions](https://support.microsoft.com/en-us/windows/zip-and-unzip-files-f6dde0a7-0fec-8294-e1d3-703ed85e7ebc)).

```{bash, include = FALSE}
unzip -o chapter-zips/r-heatmap-adv-chapt-4.zip -d chapter-zips/
```
</details>

## Exercise: Set up a project on GitHub

https://jhudatascience.org/Reproducibility_in_Cancer_Informatics/making-your-project-open-source-with-github.html

**Any feedback you have regarding this exercise is greatly appreciated; you can fill out [this form](https://forms.gle/ygSSwoGaEATA2S65A)!**

Congrats! You've started your very own project on GitHub!
We encourage you to do the same with your own code and other projects!
