
```{r, include = FALSE}
leanbuild::set_knitr_image_path()
```

# Engaging in Code Review - as a reviewer

## Learning Objectives

```{r, fig.alt="This chapter will demonstrate how to: Understand what to look for when reviewing a pull request. Provide helpful feedback in the context of code review.", out.width = "100%", echo = FALSE}
leanbuild::include_slide("https://docs.google.com/presentation/d/1IJ_uFxJud7OdIAr6p8ZOzvYs-SGDqa7g4cUHtUld03I/edit#slide=id.gfc3e8b194d_0_236")
```

## Reviewer responsibilities in code review

When reviewing a pull request, you take on responsibility to ensure that the pull request is getting the project to a better state than before.


### What to look for!

[What to look for in a code review](https://google.github.io/eng-practices/review/reviewer/looking-for.html):  

> 1. The code is well-designed.
> 2. The functionality is good for the users of the code.
> 3. Any UI changes are sensible and look good.
> 4. Any parallel programming is done safely.
> 5. The code isn’t more complex than it needs to be.
> 6. The developer isn’t implementing things they might need in the future but don’t know they need now.
> 7. Code has appropriate unit tests.
> 8. Tests are well-designed.
> 9. The developer used clear names for everything.
> 10. Comments are clear and useful, and mostly explain why instead of what.
> 11. Code is appropriately documented (generally in g3doc).
> 12. The code conforms to our style guides.

### What's useful

Identification of:  

- Functional issues - items that affect the accuracy or outcome.
- Alternate use cases where the current implementation may fail.
- Where Convention isn't being followed.  
- Where more clarity is needed.

### How to communicate it

Comments as QUESTIONS.
Comments as SUGGESTIONS.
Comments as APPRECIATIONS.

[Comments during Code Reviews](https://medium.com/@otarutunde/comments-during-code-reviews-2cb7791e1ac7)

#### Recommended reading about code review

- [On Empathy and Pull Requests](https://slack.engineering/on-empathy-pull-requests/) by @Hirpa2016.
- [Code Review Guidelines for Humans](https://phauer.com/2018/code-review-guidelines/) by @Hauer2018.
- [Your Code Sucks! – Code Review Best Practices](https://quickbirdstudios.com/blog/code-review-best-practices-guidelines/) by @Hildebr2020.
- An even longer list of [readings about code review](https://github.com/joho/awesome-code-review)
