
```{r, include = FALSE}
leanbuild::set_knitr_image_path()
```

# Using Docker to control your computing environment

## Learning Objectives

```{r, fig.alt="This chapter will demonstrate how to: Understand the goals and motivation of using Docker. Install Docker. Use basic docker commands as a part of a daily usage. Pull and use a docker image for writing code.", out.width = "100%", echo = FALSE}
leanbuild::include_slide("https://docs.google.com/presentation/d/1IJ_uFxJud7OdIAr6p8ZOzvYs-SGDqa7g4cUHtUld03I/edit#slide=id.gfa97af8537_0_81")
```

***

In the introductory part of this course, we discussed [package management using language specific package managers](https://jhudatascience.org/Reproducibility_in_Cancer_Informatics/managing-package-versions.html) like renv or conda. Recall that even if you have the same packages installed between two computers, you can still get different results!

```{r, fig.alt="Ruby has a particular computing environment she has developed her code from. This computing environment is represented as a bubble above her computer with various hexagons with version numbers as well as Rstudio and R installed on her computer. Her code ran just fine on her particular computing environment. Avi attempted to run Ruby’s code on his very different local computing environment and got an error. His computer runs the same code but came up with a different result!
", out.width = "100%", echo = FALSE}
leanbuild::include_slide("https://docs.google.com/presentation/d/1IJ_uFxJud7OdIAr6p8ZOzvYs-SGDqa7g4cUHtUld03I/edit#slide=id.gfd7f4e514e_0_256")
```

This is because package versions _do_ influence results as demonstrated by @BeaulieuJones2017.
One approach around this problem that is a bit more thorough than package managers, is using a docker container.

```{r, fig.alt="From Beaulieu-Jones and Casey S. Greene, 2017 A.) The status quo requires a reader or reviewer to find and install specific versions of dependencies. These dependencies can become difficult to find and may become incompatible with newer versions of other software packages. Different versions of packages identify different numbers of significantly differentially expressed genes from the same source code and data. B.) Containers define a computing environment that captures dependencies. In container-based systems, the results are the same regardless of the host system.", out.width = "100%", echo = FALSE}
leanbuild::include_slide("https://docs.google.com/presentation/d/1IJ_uFxJud7OdIAr6p8ZOzvYs-SGDqa7g4cUHtUld03I/edit#slide=id.gfd7f4e514e_0_0")
```
## What's Docker?

```{r, fig.alt="Ruby has a particular computing environment she has developed her code from but this time Ruby has this computing environment wrapped up in a Docker container. Avi the Associate is able to download the exact computing environment Ruby used for her analysis.
", out.width = "100%", echo = FALSE}
leanbuild::include_slide("https://docs.google.com/presentation/d/1IJ_uFxJud7OdIAr6p8ZOzvYs-SGDqa7g4cUHtUld03I/edit#slide=id.gfd7f4e514e_0_5")
```

```{r, fig.alt="Ruby and Avi are using Docker and can ensure that they are using the exact same computing environment as they are working on Ruby’s analysis.", out.width = "100%", echo = FALSE}
leanbuild::include_slide("https://docs.google.com/presentation/d/1IJ_uFxJud7OdIAr6p8ZOzvYs-SGDqa7g4cUHtUld03I/edit#slide=id.gfd7f4e514e_0_55")
```

## Install Docker


## More reading about GitHub

- [Happy Git and GitHub for the useR](https://happygitwithr.com/)
- [GitHub for data scientists](https://towardsdatascience.com/introduction-to-github-for-data-scientists-2cf8b9b25fba)
- [Intro to GitHub](https://lab.github.com/githubtraining/introduction-to-github)
- [GitHub docs about creating a Pull Request](https://docs.github.com/en/pull-requests/collaborating-with-pull-requests/proposing-changes-to-your-work-with-pull-requests/creating-a-pull-request)
- [Making a Pull Request](https://www.atlassian.com/git/tutorials/making-a-pull-request)
