
```{r, include = FALSE}
leanbuild::set_knitr_image_path()
```

# Automation as a reproducibility tool

## Learning Objectives

```{r, fig.alt="Learning Objectives This chapter will demonstrate how to: Understand how automation aids in reproducibility and efficiency. Learn the basics of GitHub Actions.", out.width = "100%", echo = FALSE}
leanbuild::include_slide("https://docs.google.com/presentation/d/1IJ_uFxJud7OdIAr6p8ZOzvYs-SGDqa7g4cUHtUld03I/edit#slide=id.g104d8c58b62_46_4")
```

***


We've discussed that a reproducible analysis can be run by someone else and obtain the same result. But what if before you bug your colleague to use their time to re-run your analysis, you had a robot re-run your analysis? Robots don't get tired or have other deadlines to respond to and can be set up to re-run your analysis at any time. This is the basis of why automation is powerful tool for reproducibility.

```{r, fig.alt="Ruby wants to know if her analysis is reproducibly so she sets up an automation tool to re-run her analysis whenever she pushes changes to her analysis. This robot has a computer for a body and says 'I will re-run Ruby’s analysis consistently and instantly upon whatever trigger she sets up.''", out.width = "100%", echo = FALSE}
leanbuild::include_slide("https://docs.google.com/presentation/d/1IJ_uFxJud7OdIAr6p8ZOzvYs-SGDqa7g4cUHtUld03I/edit#slide=id.g104d8c58b62_46_9")
```
